---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hmer

<!-- badges: start -->
<!-- badges: end -->

## Overview

The goal of hmer is to make the process of history matching and emulation accessible and easily useable by modellers, particularly in epidemiology. The central object of the process is an `Emulator`: a statistical approximator for the output of a complex (and often expensive) model that, given a relatively small number of model evaluations, can give predictions of the model output at unseen points with the appropriate uncertainty built-in. Using these we may follow a process of 'history matching', where unfeasible parts of the parameter space are ruled out. Sampling parameter sets from the remaining region allows us to train more accurate emulators, which allow us to remove more of the space, and so on. The hmer package contains tools for the automated construction of emulators, visualisations for diagnostic checks and exploration of parameter space, and a means by which new points can be proposed.

## Installation

You can install the development version of hmer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andy-iskauskas/hmer")
```

## Example

The core functions of the package are detailed below, using built-in toy data.


```{r example}
library(hmer)
ems <- emulator_from_data(input_data = SIRSample$training,
                          output_names = names(SIREmulators$targets),
                          ranges = list(aSI = c(0.1, 0.8), aIR = c(0, 0.5), aSR = c(0, 0.05)))
validation <- validation_diagnostics(ems, SIREmulators$targets, SIRSample$validation, plt = FALSE)
new_points <- generate_new_runs(ems, 50, SIREmulators$targets)
```

## Learning Emulation and History Matching

There is a wealth of published information on Bayes Linear emulation, history matching, and the more general framework of uncertainty quantification; the easiest way to learn how to use the hmer package, however, is to look through the vignettes within. The "Low Dimensional Examples" vignette is a good place to start.

```{r vignettes, eval = FALSE}
browseVignettes("hmer")
vignette("low-dimensional-examples", package = 'hmer')
```

